---
title: "p8130_hw5_ps3194"
author: "Pangsibo Shen"
date: "11/19/2020"
output: html_document
---

```{r message=FALSE}
library(tidyverse)
```

## Problem 1

Given the non-normal distributions, now you are asked to use an alternative, non-parametric test
to assess and comment on the difference in Ig-M levels between the two groups (please ignore
unanswered and missing values). You can use R to perform the calculations, but make sure you
state test that you used, the hypotheses, test statistic, p-value and provide interpretation in the
context of the problem.

```{r message=FALSE}
#load antibodies dataset
antibodies = read_csv("./data/Antibodies.csv") %>%
  drop_na()

#extract IgM values for Normal group
Normal = antibodies %>%
  filter(Smell == "Normal") %>%
  pull(Antibody_IgM)
#extract IgM values for Altered group
Altered = antibodies %>%
  filter(Smell == "Altered") %>%
  pull(Antibody_IgM)
```
```{r message=FALSE}
#Graphic methods to test for normality
par(mfrow=c(2,2))
hist(Normal, xlab="Antibody Ig-M levels", freq=T, main="Normal Group")
hist(Altered, xlab="Antibody Ig-M levels", freq=T, main="Altered Group")
qqnorm(Normal, col=2, pch=19, cex=1.5)
qqline(Normal, col = 1,lwd=2,lty=2)

qqnorm(Altered, col=2, pch=19, cex=1.5)
qqline(Altered, col = 1,lwd=2,lty=2)
```


```{r message=FALSE}
# Non-parametric Wilcoxon-Rank Sum test: two-independent groups
wilcox.test(Normal, Altered, mu=0)
```

